---
title: 'Reproducible Research: Assignment 1'
output:
  html_document:
  keep_md: yes
  pdf_document: default
---

This assignment is about analizing and showing interesting imformation on human activity.   
The file activity.csv is given. This file consists of 15768 rows. each row has 3 columns such as steps,date, and interval.  
interval has value from 0,5,10,15 and so on until 2355. 0 means activity from 0 minute to 5 minute of a day.  5 means activity from 5 to 10 of a day and so on.   


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading and preprocessing the data
First of all, I loaded csv file using read.csv. After that, some rows are showed. 

```{r}
d <- read.csv('activity.csv')
```

##basic structure
It shows basic structure of this data file.
```{r}
head(d)
summary(d)
nrow(d)
```


# What is mean total number of steps taken per day?
ddply is used to get the number of total steps for each day.
mean value can be gotten through dividing the number of stpes by the number of days.  
median value can be showed by getting median value after sorting.  

median/mean functions are used for getting values simply.  

```{r}
library(plyr)
out <- ddply(d, c("date"),summarize, steps=sum(steps))

mean_val <-mean(out$steps, na.rm = T)
median_val <- median(out$steps, na.rm=T)

mean_val
median_val 
head(out)
```


##histogram
The output graph is saved in plot0.png file.

```{r scatterplot0,fig.height=4}
hist(out$steps ,breaks=100, main=" total number of steps taken per day",
     xlab="steps", ylab="frequency", col="pink")
dev.copy(png,'plot0.png')
dev.off()
```


# What is the average daily activity pattern?
1 day has 24 hours. 1 hour has 60 minutes. Therefore 24*60 means minutes of a day.  
24*60 is divided by 5 because they made data  every 5 minutes.
As a result, total number of interval count is 288 (24*60/5).  
"for loop"" is used to get average values for each time interval.  
For example, mean(first row of d + 289th row + 577th row ..... 16993th row + 17281th row)   is the average steps value of first interval.   

out2 is used for saving the mean value for each interval. 

```{r}
days <- nrow(out)
interval_cnt <- 24*60/5   

out2 <- rep(0, interval_cnt)

for(i in 1:interval_cnt)
{
    some <- seq(i, nrow(d), by= interval_cnt)
    out2[i] <- mean( d[some,]$steps, na.rm=T)
}
```

##Plot it.
```{r scatterplot1,fig.height=4}
plot(out2 , type="l", main="average daily activity pattern",
     xlab="5 minutes interval",ylab="steps", col="red" , lwd=3)
dev.copy(png,'plot1.png')
dev.off()
```

##maximum active interval value

```{r}
max_interval <- max(out2)
print(max_interval)
```


# Imputing missing values
set the missing value to the average. out2 is the daily activity pattern.   
As calculated before, 288(24*60/5) is interval count of a day.
```{r}
d_imputed <- d
for(i in 1:nrow(d_imputed) )
{
    if( is.na(d_imputed[i,1])  == TRUE)
    {
       d_imputed[i,1] <- out2[ i %% 288 + 1]

    }
}

out4hist <- ddply(d_imputed, c("date"),summarize, steps=sum(steps))
```

##histogram
```{r scatterplot2,fig.height=4}
hist(out4hist$steps ,breaks=100, 
     main="total number of steps taken per day after IMPUTING" , xlab="steps",ylab="frequency", col="magenta")
dev.copy(png,'plot2.png')
dev.off()
```

# Are there differences in activity patterns between weekdays and weekends?
As we expected, peopel tend to move more on the weekend.
I'll calculate the average value of steps on the weekdays/weekends.   

I found timeDate package after struggling from identifying weekend.   


weekdays_n : the number of weekdays  
weekdays_sum: total number of steps on weekdays.  

weekend_n : the number of weekend.    
weekend_sum: total number of steps on weekend.



```{r}
#install.packages("timeDate")
library(timeDate)

out3 <- ddply(d_imputed, c("date"),summarize, steps=sum(steps))


weekdays_n <- 0
weekdays_sum <-0

weekend_n <- 0
weekend_sum <- 0


for(i in 1:nrow(out3))
{
    if( as.logical(isWeekend(as.Date(out3[i,1]), wday = 1:5)) )
    {
       weekend_n <- weekend_n +1 
       weekend_sum <- weekend_sum + out3[i,2]
    
    }
    else
    {
       weekdays_n <- weekdays_n + 1
       weekdays_sum <- weekdays_sum + out3[i,2]
    
    }
}
```



##print average steps in weekend & weekdays
As expected, people walk more on weekends  compared to weekdays.
```{r}
print( weekend_sum / weekend_n )
print( weekdays_sum / weekdays_n )
```





## weekdays vs weekends with plotting

24*60/5 means the number of interval count per a day. I calculated above.  

weekdays_arr: it keep data about weekdays for plotting  
weekends_arr: it keep data about weekends for plotting  
if it is record for weekends, the value is adding to weekends_arr using idx value.

idx:  index for weekdays_arr, weekends_arr 

```{r}

interval_cnt <- 24*60/5   
weekdays_arr<- rep(0, interval_cnt)
weekends_arr <- rep(0,interval_cnt)

for(i in 1:nrow(d_imputed) )
{
    steps    = d_imputed[i,1]
    date     = d_imputed[i,2]
    idx      = d_imputed[i,3]/5        
    
    if( as.logical(isWeekend(as.Date(date), wday = 1:5)) )
    {
        weekdays_arr[idx] <- weekdays_arr[idx] + steps
  
    }
    else
    {
        weekends_arr[idx] <- weekends_arr[idx] + steps
    }

}

```


###weekdays
```{r}


   
plot(weekdays_arr , type="l", main="weekdays activity pattern",
     xlab="5 minutes interval",ylab="steps", col="blue" , lwd=3)
dev.copy(png,'plot3.png')
dev.off()
```

###weekends
```{r}
plot(weekends_arr , type="l", main="weekends activity pattern",
     xlab="5 minutes interval",ylab="steps", col="blue" , lwd=3)
dev.copy(png,'plot4.png')
dev.off()
```



